---
title: "Melanoma mIF: Figure 4"
subtitle: "Adapted from Code by Xiyu Peng, PhD"
editor: source
date: "Last Updated: `r format(Sys.Date(), format = '%B %d, %Y')`"
embed-resources: TRUE
execute: 
  echo: FALSE
  message: FALSE
  warning: FALSE
format:
  html:
    toc: TRUE
    toc-location: left
    toc-depth: 3
    page-layout: full
---

```{r packages, include=FALSE}
library(tidyverse)
library(ggplot2)
library(patchwork)
```

```{r import, include=FALSE}
# Set gtsummary theme
theme_gtsummary_compact()

# Import data `df_dist`
```

```{r}
# Parameters for plots
nbins = 72
colors = 
  list("3:Bcell" = "#007CBA",
       "5:CD8T" = "#FFC20E",
       "4:CD4T" = "#4C8B2B",
       "6:Treg" = "#8FC7E8",
       "2:Macrophage" = "#83276B",
       "1:tumor/epi" = "red",
       "7:other" = "grey"
  )
new_labels =
  list("3:Bcell" = "B cell",
       "5:CD8T" = "CD8 T cell",
       "4:CD4T" = "CD3+CD8- T cell",
       "6:Treg" = "Treg",
       "2:Macrophage" = "Macrophage",
       "1:tumor/epi" = "Tumor/Epithelial",
       "7:other" = "Other")
```

# Included Samples

These analyses are limited to non-LN samples that have B cells (N = `r df_dist %>% distinct(tissue_id) %>% nrow()`). There are `r df_dist %>% filter(response_col == "Yes") %>% distinct(tissue_id) %>% nrow()` responders and `r df_dist %>% filter(response_col == "No") %>% distinct(tissue_id) %>% nrow()` non-responders.

# Figure 4a

```{r}
# Randomly sample 1000 cells
set.seed(123)
df_dist_sub =
  df_dist %>% 
  group_by(tissue_id) %>%
  slice_sample(n = 1000) %>% 
  ungroup()
```

Here, 1000 cells are randomly sampled from each tissue (N = `r df_dist_sub %>% distinct(tissue_id) %>% nrow()` samples, N = `r df_dist_sub %>% nrow()` cells).

Note that some samples have fewer than 1000 cells, in which case all cells are taken.

```{r, fig.height = 4, fig.width = 6}
# Plotting function
plot_bar_percentage = function(x, celltype) {
   
  df_input = data.frame(x = x, celltype = celltype)
   
  df =
    df_input %>% 
    mutate(cut_x = cut(x, nbins)) %>% 
    group_by(cut_x, celltype) %>% 
    reframe(n = length(cut_x)) %>% 
    ungroup() %>% 
    group_by(cut_x) %>% 
    reframe(celltype, perc = n/sum(n)*100)
  
  df_0_label =
    tibble(
      cut_x = as.factor("0"),
      celltype = df %>% ungroup() %>% distinct(celltype) %>% pull(celltype),
      perc = 0
    ) 
  
  df_plot = 
    rbind(df, df_0_label) %>% 
    mutate(cut_x = fct_relevel(cut_x, "0")) %>% 
    arrange(cut_x)

  x_breaks = df_plot %>% distinct(cut_x) %>% pull(cut_x)
  x_breaks = x_breaks[c(1, seq(from = 15, by = 14, length.out = 5))]
  x_labels = as.character(seq(0, 250, 50))
  
  df_plot %>% 
    ggplot(., aes(x = cut_x, y = perc, fill = celltype, color = celltype)) +
    geom_bar(stat = "identity", position = "stack") +
    scale_x_discrete(name = "Distance to the closest B cell (µm)", breaks = x_breaks, labels = x_labels) +
    ylab("% immune cells") +
    theme_classic() +
    scale_fill_manual(name = "Major cell type", values = colors, labels = new_labels) +
    scale_colour_manual(guide = "none", values = colors) +
    theme(legend.position = "bottom")
}

# Generate plot
plot_bar_percentage(df_dist_sub$micron, df_dist_sub$majortype)
```

# Figure 4c

```{r}
# Cells from responders
df_dist_sub_responders = 
  df_dist_sub %>% 
  filter(response_col == "Yes")

# Cells from non-responders
df_dist_sub_nonresponders = 
  df_dist_sub %>% 
  filter(response_col == "No")
```

The same data used for `Figure 4a` is split by response (N = `r df_dist_sub_responders %>% distinct(tissue_id) %>% nrow()` responders with N = `r df_dist_sub_responders %>% nrow()` cells, N = `r df_dist_sub_nonresponders %>% distinct(tissue_id) %>% nrow()` non-responders with N = `r df_dist_sub_nonresponders %>% nrow()` cells).

```{r}
# Generate plots
p_responders = plot_bar_percentage(df_dist_sub_responders$micron, df_dist_sub_responders$majortype) + ggtitle("Responders")

p_nonresponders = plot_bar_percentage(df_dist_sub_nonresponders$micron, df_dist_sub_nonresponders$majortype) + ggtitle("Non-responders")
```

```{r, fig.width = 12, fig.height = 4}
# Print plots
wrap_plots(p_responders, p_nonresponders, guides = "collect") & theme(legend.position = "bottom")
```

# Figure 4d

```{r}
df_dist_sub_tcell =
  df_dist_sub %>% 
  filter(majortype %in% c("5:CD8T", "4:CD4T", "6:Treg"))
```

1000 cells randomly sampled from each tissue (same as data used for Figure 4a), then filtered to T cells. N = `r df_dist_sub_tcell %>% distinct(tissue_id) %>% nrow()` samples, N = `r df_dist_sub_tcell %>% nrow()` cells (not all samples had T cells sampled).

```{r}  
p_4d_1 = 
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron, fill = Classify_TCF1.7)) + 
  geom_histogram(position = "stack", bins = nbins) + 
  labs(x = "Distance to the closest B cell (µm)", y = "Cell count") + 
  theme_classic() +
  scale_fill_manual(
    name = "TCF1/7",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 1000)

p_4d_2 = 
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron, fill = Classify_LAG3)) +
  geom_histogram(position = "stack", bins = nbins) +
  labs(x = "Distance to the closest B cell (µm)", y = "Cell count") +
  theme_classic() +
  scale_fill_manual(
    name = "LAG-3",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 1000)

p_4d_3 =
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron,fill = Classify_Ki67)) + 
  geom_histogram(position = "stack", bins = nbins) +
  labs(x = "Distance to the closest B cell (µm)", y = "Cell count") +
  theme_classic() +
  scale_fill_manual(
    name = "Ki67",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 1000)
```

```{r,fig.width = 12, fig.height = 3}
wrap_plots(p_4d_1, p_4d_2, p_4d_3, nrow = 1)
```

# Figure 4e

1000 cells randomly sampled from each tissue (same as data used for Figure 4a), then filtered to T cells. N = `r df_dist_sub_tcell %>% distinct(tissue_id) %>% nrow()` samples, N = `r df_dist_sub_tcell %>% nrow()` cells (not all samples had T cells sampled).

```{r}  
p_4e_1 = 
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron, color = Classify_TCF1.7)) + 
  geom_density() + 
  labs(x = "Distance to the closest B cell (µm)", y = "Density") + 
  theme_classic() +
  scale_color_manual(
    name = "TCF1/7",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 0.02)

p_4e_2 = 
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron, color = Classify_LAG3)) +
  geom_density() +
  labs(x = "Distance to the closest B cell (µm)", y = "Density") +
  theme_classic() +
  scale_color_manual(
    name = "LAG-3",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 0.02)

p_4e_3 =
  df_dist_sub_tcell %>% 
  ggplot(aes(x = micron,color = Classify_Ki67)) + 
  geom_density() +
  labs(x = "Distance to the closest B cell (µm)", y = "Density") +
  theme_classic() +
  scale_color_manual(
    name = "Ki67",
    values = c("#67A9CF","#EF8A62")) +
  theme(legend.position = "bottom") +
  ylim(0, 0.02)
```

```{r,fig.width = 12, fig.height = 3}
wrap_plots(p_4e_1, p_4e_2, p_4e_3, nrow = 1)
```

For each comparison, generate 1000 Kolmogorov-Smirnov p-values by repeatedly randomly sampling 1000 cells. Calculate median p-values and adjust these median p-values using the Bonferroni method.

```{r}
p_values_tcf17 = c()
p_values_lag3 = c()
p_values_ki67 = c()

set.seed(123)

for (i in 1:1000) {
  
  df_sample =
    df_dist_sub_tcell %>% 
    group_by(tissue_id) %>% 
    slice_sample(n = 1000) %>% 
    ungroup()
  
  p_values_tcf17 = 
    c(p_values_tcf17, 
      ks.test(df_sample$dist2Bcell[df_sample$Classify_TCF1.7 == "Negative"], df_sample$dist2Bcell[df_sample$Classify_TCF1.7 == "Positive"])$p.value)
  
  p_values_lag3 = 
    c(p_values_lag3, 
      ks.test(df_sample$dist2Bcell[df_sample$Classify_LAG3 == "Negative"], df_sample$dist2Bcell[df_sample$Classify_LAG3 == "Positive"])$p.value)
  
  p_values_ki67 = 
    c(p_values_ki67, 
      ks.test(df_sample$dist2Bcell[df_sample$Classify_Ki67 == "Negative"], df_sample$dist2Bcell[df_sample$Classify_Ki67 == "Positive"])$p.value)

}

df_ks_results =
  tibble(
    marker = c("TCF1/7", "LAG3", "Ki67"),
    median_p = c(median(p_values_tcf17), median(p_values_lag3), median(p_values_ki67))
  ) %>% 
  mutate(
    median_p_adjusted = p.adjust(median_p, method = "bonferroni"),
    across(c(median_p, median_p_adjusted), ~ ggsurvfit::format_p(.))
  )

df_ks_results %>% 
  knitr::kable()
```
